---
title: "Cars_mpg_report"
author: "Matthias Booten"
date: "26-12-2019"
output: html_document
---

```{r intro, echo=FALSE, message=FALSE, warning=FALSE}

#Importing the file and cleaning it
                           
library(readr)
mydf <- read_delim("C:\\Users\\matth\\Desktop\\Data Analyst\\R\\PrepMaterial\\auto_mpg.csv", ";", 
                       escape_double = FALSE, trim_ws = TRUE) #telling R to read CSV files for file upload 
#bView(mydf) # importing CSV file into R environment and naming it "mydf"

#Deleting rows with missing values

library(tidyr)              # calling a package to use drop_na function

mydf<-drop_na(mydf)  

#Deleting a select column from the dataset

library(dplyr)                   # calling a package to use the function to drop select columns

mydf = select(mydf, -orogin)     # removing the "orogin" column from mydf

```

## Insights gained from performing statistics on the cars dataset

This files includes the general business insights derived from the auto mpg dataset. The goal is gaining insights in terms of fuel economy. We can conclude that the most efficient
quartile of cars start at a fuel economy of 29mpg while average fuel economy remains at 23.45 mpg.

```{r mpg Business Questions, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

### Business Questions ###

#What is the 75th percentile of mpg metric for available cars?
summary(mydf$mpg)   # The 75th percentile of cars by mpg are between 29 and 46



# What are the car names that match the 75th percentile of cars by mpg?
good_cars<-subset(mydf, mydf$mpg>=29)  # Creating a new data frame with cars that have mpg of 29 or higher 

# What cars within the 75th percentile of cars by mpg are built in 1975 or later?
new_cars<-subset(good_cars, good_cars$model_year>=75)   # Creating a new object to see what cars are made in 1975 or lower

# How did the vehicle fuel consumption improve between 1970 and 1982 by 4-year increments?

# Using a for loop to create 3 groups by model_year (70-73, 74-77, 78-81)
for(i in 1:nrow(mydf)){
  if(mydf$model_year[i]>=70 & mydf$model_year[i]<=73){mydf$model_year_group[i]<-'group #1'}
  else if(mydf$model_year[i]>=74 & mydf$model_year[i]<=77){mydf$model_year_group[i]<-'group #2'}
  else if(mydf$model_year[i]>=78 & mydf$model_year[i]<=81){mydf$model_year_group[i]<- 'group #3'}
  else if(mydf$model_year[i]==82){mydf$model_year_group[i]<- 'group #4'}
}

```

The next part of this R report looks into the relation between fuel efficiency and multiple characteristical parameters of our vehicles. With the success being defined as a car belonging to the vehicles with an mpg of 30 or higher, the logistic regression proved that: 

 with every increase of cylinders by 1 unit, the probability of success decrease by 5.56%
 with every increase of horsepower by 1 unit, the probability of success decrease by 10.24%
 with every increase of weight by 1 unit, the probability of success decrease by 0.09%
 with every increase of acceleration by 1 unit, the probability of success decrease by   23.48%

```{r mpg Business Questions2,echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
####################################################################???
#Now we want to find out about the fuel consumption <30MPG is success

#Build a logistic regression for mpg
#Create a column "good_bad" that represents success with 1 being success and 0 = failure

for(i in 1:nrow(mydf)){
  if (mydf$mpg[i]>=30){mydf$good_bad[i]<-"1"}
  else{mydf$good_bad[i]<-"0"}
}

#converting the good_bad column back to nummeric for further use
mydf$good_bad <- as.numeric(mydf$good_bad)


#Make a logistic regression to see the related factors

mpg_influence<-glm(good_bad ~ cylinders+horsepower+weight+acceleration, data=mydf, family="binomial")
summary(mpg_influence)

```
Following decision tree helps in identifying vehicles having a fuel economy of 30 mpg or higher.

In the dataset, 302 cars travel up to 29 miles per gallown, and 90 cars travel 30 or more miles per gallon.

For instance: From the cars that have 76 bhp or less, 71 travel 30 miles or more per gallon.

The graph below shows us the true false prediction accuracy of the  logistic regression  (red) VS the decision tree (black). We can see that the logistic regression has a bigger surface area under the curve. This indicates that the logistic regression is the most reliable to base business decisions on.


```{r mpg Business Questions3, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

library(rpart)
library(rpart.plot)
tree_mydf<-rpart(good_bad~cylinders+horsepower+weight+acceleration, data=mydf, method="class")
rpart.plot(tree_mydf, extra=1, type=1)   

tree_pred<-predict(tree_mydf, mydf, type="prob")              #prediction model for tree
logit_pred<-predict(mpg_influence, mydf, type="response")     #prediction model for the logistic regression

library(ROCR)
pred_val_tree<-prediction(tree_pred[,2], mydf$good_bad)
pred_val_logit<-prediction(logit_pred, mydf$good_bad)

perf_tree<-performance(pred_val_tree, "tpr", "fpr")
perf_logit<-performance(pred_val_logit, "tpr", "fpr")
plot(perf_tree, col="black")
plot(perf_logit, col='red', add=TRUE)


```

## Including Plots

Creates plots with 4 groups with years between 1970-1973,1974-1977,1978-1981 and 1982 respectively. We can see that over the years mpg increased.

The overview of car groups 1 through 4 are plotted

```{r plotting, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

library(ggplot2)    # Importing ggplot2 package to visualize the data

group_mpg_plot<-ggplot(mydf, aes(mydf$model_year_group, mydf$mpg)) #plotting the relationship between group number and mpg
group_mpg_plot + geom_point(shape=3)
group_mpg_plot + geom_point(aes(colour=model_year_group)) #adding color to separate groups

```


